---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
data<-read.csv("~/Documents/activity.csv", stringsAsFactor = FALSE)
str(data)
```

As you would notice the date variable is character. We would transform it into a date type

```{r echo=TRUE}
data$date<-as.Date(data$date, format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r echo=TRUE}

library(dplyr)
data%>%
  group_by(date)%>%
  summarise(Total_Steps=sum(steps,na.rm=TRUE))
```

2. Histogram of the total number of steps taken each day

```{r echo=TRUE}
library(ggplot2)
data%>%
  group_by(date)%>%
  summarise(Total_Steps=sum(steps,na.rm=TRUE))%>%
  ggplot(aes(Total_Steps)) + geom_histogram(binwidth=1000) + xlim(1000,18000) + 
  ggtitle("Total Number of Steps per day")
```

3. Mean and Median of the total number of steps taken per day

```{r echo=TRUE}
stepdata<-data%>%
  group_by(date)%>%
  summarise(Total_Steps=sum(steps,na.rm=TRUE))
```  
Mean Total Number of steps
```{r echo=TRUE}
mean(stepdata$Total_Steps)
```

Median Total Number of Steps
```{r echo=TRUE}
median(stepdata$Total_Steps)
```


## What is the average daily activity pattern?

1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken

```{r echo=TRUE}
timeseries<-data%>%
  group_by(interval)%>%
  summarise(Average=(sum(steps, na.rm=TRUE)/61)) #61 is the total number of days between Oct and November

plot(x=timeseries$interval, y=timeseries$Average, type="l", xlab="Interval", ylab="Average number of steps")
```

2. The 5-minute interval that contains the maximum number of steps on average across all the days in the dataset, 

```{r echo=TRUE}
timeseries<-data%>%
  group_by(interval)%>%
  summarise(Average=(sum(steps, na.rm=TRUE)/61))%>%
  arrange(desc(Average))

timeseries[1,1]
```

## Imputing missing values

1. Total number of missing values in the dataset
```{r echo=TRUE}
length(which(is.na(data)))
```

2. New dataset with missing values filled with mean for that time interval

```{r echo=TRUE}
newdata<-data%>%
  group_by(interval)%>%
  mutate(steps=ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
  
newstepdata<-newdata%>%
  group_by(date)%>%
  summarise(total=sum(steps, na.rm=TRUE))
  
ggplot(data=newstepdata, aes(total))+geom_histogram(binwidth=1000)+ xlim(1000,18000)

mean(newstepdata$total)
median(newstepdata$total)
```

## Are there differences in activity patterns between weekdays and weekends?

``` {r echo=TRUE}

newdata$day<-ifelse(weekdays(newdata$date) %in% c("Saturday","Sunday"), "weekend", "weekday")

library(ggplot2)
newdata%>%
  group_by(interval,day)%>%
  summarise(avg_steps=mean(steps))%>%
  ggplot(aes(x=interval, y=avg_steps)) + geom_line()+facet_grid(day~.) + xlab("Interval")+ylab("Average Steps")
